---
## "Reproducible Research: Peer Assessment 1"

---

### **Loading and preprocessing the data**
Load Libraries ddplyr and ggplot2.

```{r Load libraries}
library("plyr")
library("ggplot2")
```

Unzip the activity.zip and read the activity.csv file
```{r Unzip and read.csv}
# Unzip the activit.zip file
unzip("activity.zip")
#Read activity.csv file
activity <- read.csv("activity.csv", header=T)
```

The class for Activity$date must be date instead of the current class factor
```{r Set the calss of date}
class(activity$date)
# Class of date is a factor needs to be date.
activityDate <-transform(activity, date = as.Date(date))
class(activityDate$date)
# Class date is Date
```

## **What is mean total number of steps taken per day?**

before we can calculate the number of steps per day we must
set the class for Activity$date to be date instead of the current class factor
```{r Set the calss of date}
class(activity$date)
# Class of date is a factor needs to be date.
activityDate <-transform(activity, date = as.Date(date))
class(activityDate$date)
# Class date is Date
```

now we can calculate the total number of steps per day.
```{r mean total number of steps per day}
MstepsPerDay <- ddply(activity, ~date, summarise, steps = sum(steps))
```

In order to show the number of steps per day I will show the result of the stepsPerDay in a histogram
```{r Histogram mean tot. steps per day }
p <- ggplot (MstepsPerDay, aes (steps))
p <- p + geom_histogram(fill = "yellow", color = "black") + theme_bw(base_family = "Verdana")
p + ggtitle ("Total number of steps per day")
p + labs(x = "Step per day", y = "Number of times it occured")
```


## What is the average daily activity pattern?



## Imputing missing values



## Are there differences in activity patterns between weekdays and weekends?
